What's in a Name? Linear Temporal Logic Literally Represents Time Lines

### 2023-01-02 paper/main.tex
fix typo
minor change

    \begin{highlights}
        \item We transform Linear Temporal Logic (LTL) formulas to timeline visualizations for the set of satisfying traces.
        \item We translate LTL formulas to \Buchi automata, then to $\omega$-regular expressions, then simplify them and use graph visualization tools to draw timelines.
        \item Our algorithm is sound and complete.
        \item Our tool, \tool, effectively constructs small timeline visualizations for a range of LTL formulas.
    \end{highlights}

@@

\begin{definition}[Linear Temporal Logic (LTL)]
    % cite Kristin's survey paper
    The syntax of an LTL formula over a set of atomic propositions $\AP$, where $p\in\AP$ is a propositional variable, is given by the following grammar:
    \begin{align*}
        \Phi ::= p \mid \neg \Phi \mid \Phi \land \Phi \mid \Phi \lor \Phi \mid \Phi \limplies \Phi \mid \always \Phi \mid \eventually \Phi \mid \nextt \Phi \mid \Phi \stronguntil \Phi \mid \Phi \weakrelease \Phi
    \end{align*}\label{ltl-defn}
\end{definition}

@@

%KYR: Here we nicely show (with pictures as much as possible) all of the individual statistics and results we gathered (note aggregation occurs below).
\paragraph{Real-world datasets} We study the feasibility of our tool on a benchmark suite of LTL formulas gathered from real world use cases.

@@

In \cref{fig:scalability}, the x-axis ranges from $1$ to $6$, after which point it hits the $30$ seconds timeout we set, as the timeline length grows exponentially fast.

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{img/scale.png}
    \caption{Scalability measured with binary counter examples from Rozier and Vardi \cite{RV10}}
    \label{fig:scalability}
\end{figure}

@@

\section{Conclusion}
The Achilles heel of formal verification is specification; formal methods are only as effective at verification as their specifications are at describing the essential properties to verify. Yet, specification remains the biggest bottleneck to the use of formal methods \cite{Roz16}. LTL is one of the most popular specification logics for industrial-scale critical systems; in the space domain alone, it is currently encapsulating specifications for the development of the NASA Lunar Gateway \cite{DBR21}, the Air Force Research Laboratory/Collins Aerospace Spacecraft Collision Avoidance system \cite{HDWF21}, Space Systems Finland's Attitude and Orbit Control Systems (AOCS) \cite{ILLTV13}, and NASA/JPL's Europa Lander Mission Concept \cite{CDRWRWL22} just to name a few. 
 %
Yet the humans that need to use formal verification tools and deeply understand their results struggle to validate that LTL formulas capture the specifications they are meant to capture. A major contributor to LTL's popularity is the propensity of humans to think of requirements in terms of timelines. This inspired the creation of LTL in the first place, as a logic that ``intuitively'' represents timelines. Our work serves to reinforce that connection, enabling visualization of most realistic LTL formulas as timelines. By releasing \tool, we contribute to better validation capabilities for LTL specifications and aid the effort to make formal verification more accessible and wide-spread. 
%We hope that this will assist system designers in formulating error-free specifications and thus improve the effectiveness of verification tools for \ltl.

Future extensions of this work include optimizations to the algorithm and implementation to improve performance and scalability. While we have chosen visual elements that succinctly represent timelines, it would be informative to conduct a study on different possible visualizations and which of the many ways of representing different timeline elements humans find most intuitive. It is possible that factors of context, such as the type of requirement an LTL formula describes, change its optimal timeline representation.

>>>>

    \begin{highlights}
        \item We transform Linear Temporal Logic (\ltl) formulas to timeline visualizations for the set of satisfying traces.
        \item We translate \ltl\ formulas to \Buchi automata, then to $\omega$-regular expressions, then simplify them and use graph visualization tools to draw timelines.
        \item Our algorithm is sound and complete.
        \item Our tool, \tool, effectively constructs small timeline visualizations for a range of \ltl\ formulas.
    \end{highlights}

@@

\begin{definition}[Linear Temporal Logic (LTL)]
    % cite Kristin's survey paper
    The syntax of an \ltl\ formula over a set of atomic propositions $\AP$, where $p\in\AP$ is a propositional variable, is given by the following grammar:
    \begin{align*}
        \Phi ::= p \mid \neg \Phi \mid \Phi \land \Phi \mid \Phi \lor \Phi \mid \Phi \limplies \Phi \mid \always \Phi \mid \eventually \Phi \mid \nextt \Phi \mid \Phi \stronguntil \Phi \mid \Phi \weakrelease \Phi
    \end{align*}\label{ltl-defn}
\end{definition}

@@

%KYR: Here we nicely show (with pictures as much as possible) all of the individual statistics and results we gathered (note aggregation occurs below).
\paragraph{Real-world datasets} We study the feasibility of our tool on a benchmark suite of \ltl\ formulas gathered from real world use cases.

@@

In \cref{fig:scalability}, the $x$-axis ranges from $1$ to $6$, after which point it hits the $30$ seconds timeout we set, as the timeline length grows exponentially fast.

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.6]{img/scale.png}
    \caption{Scalability measured with binary counter examples from Rozier and Vardi \cite{RV10}}
    \label{fig:scalability}
\end{figure}

@@

\section{Conclusion}
The Achilles heel of formal verification is specification; formal methods are only as effective at verification as their specifications are at describing the essential properties to verify. Yet, specification remains the biggest bottleneck to the use of formal methods \cite{Roz16}. \ltl\ is one of the most popular specification logics for industrial-scale critical systems; in the space domain alone, it is currently encapsulating specifications for the development of the NASA Lunar Gateway \cite{DBR21}, the Air Force Research Laboratory/Collins Aerospace Spacecraft Collision Avoidance system \cite{HDWF21}, Space Systems Finland's Attitude and Orbit Control Systems (AOCS) \cite{ILLTV13}, and NASA/JPL's Europa Lander Mission Concept \cite{CDRWRWL22} just to name a few.
 %
Yet the humans that need to use formal verification tools and deeply understand their results struggle to validate that \ltl\ formulas capture the specifications they are meant to capture. A major contributor to \ltl's popularity is the propensity of humans to think of requirements in terms of timelines. This inspired the creation of \ltl\ in the first place, as a logic that ``intuitively'' represents timelines. Our work serves to reinforce that connection, enabling visualization of most realistic \ltl\ formulas as timelines. By releasing \tool, we contribute to better validation capabilities for \ltl\ specifications and aid the effort to make formal verification more accessible and wide-spread.
%We hope that this will assist system designers in formulating error-free specifications and thus improve the effectiveness of verification tools for \ltl.

Future extensions of this work include optimizations to the algorithm and implementation to improve performance and scalability. While we have chosen visual elements that succinctly represent timelines, it would be informative to conduct a study on different possible visualizations and which of the many ways of representing different timeline elements humans find most intuitive. It is possible that factors of context, such as the type of requirement an \ltl\ formula describes, change its optimal timeline representation.
